---
title: "Decontamination Vignette"
author: "Rachel Griffard"
date: "`r Sys.Date()`"
output: html_document
---

# Load library

```{r, message=FALSE, warning=FALSE}
library(PACKAGE)
```

## Read in data and format

Read in the sample 16S rRNA sequencing data

<https://www.ebi.ac.uk/ena/browser/view/PRJEB37663>

```{r}

```

Users can format the data in a *phyloseq* or *SummarizedExperiment* object for use within PACKAGE.

## Decontamination

Users input data into the single **decontaminate** function. The returned object will contain only those features that passed the deontamination pipeline.

```{r}
decontaminate(object, contaminant = 2, prevalence_threshold = 0.5, blocklist = blocklist, venn = FALSE, dashboard = FALSE)
```

Within this function, four separate methods to identify potential contaminants within the data:

1.  Identify features with different abundances between batches

2.  Identify features higher in negative control and lower in samples

3.  Identify features with different abundance in different batches for technical replicates

4.  Identify previously known contaminants\
    Users can input a character list with known contaminants or use the automatic list included within PACKAGE.
